---
title: "Assignment 1"
author: "Hans"
date: "9/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Single analysis
phasePlaneAnalysis(model, xlim = xlim, ylim = ylim, parameters = pars)

# Define the model and plot flow field, null clines and one trajectory
```{r define_model}
library("phaseR")

# define 2D ODE model

model = function(timepoint, state, parameters) { 
  with(as.list(c(state, parameters)), {
    dx = -2 * x - 3 * y 
    dy = -x - 2 * y 
    list(c(dx, dy))
  }) 
}

# for setting the axis limits
xlim = c(-10, 10) 
ylim = c(-10, 10)

# no parameters in the ODE, so we set pars as an empty vector
pars = vector()
# draw vector field
model.flowField = flowField(xlab = "x", ylab = "y", model, 
                            xlim = xlim, ylim = ylim, 
                            parameters = pars, 
                            points = 19, add = FALSE, 
                            state.names = c("x", "y"))

# draw nullclines
model.nullclines = nullclines(model, xlim = xlim, ylim = ylim,
                              parameters = pars,
                              points = 500, state.names = c("x", "y"))

# draw trajectory
model.trajectory = trajectory(model, y0 = c(-5, -5), tlim = c(0, 10), 
                              parameters = pars, state.names = c("x", "y"))
model.trajectory = trajectory(model, y0 = c(0, -10), tlim = c(0, 10), 
                              parameters = pars, state.names = c("x", "y"))
model.trajectory = trajectory(model, y0 = c(-5, 5), tlim = c(0, 10), 
                              parameters = pars, state.names = c("x", "y"))
model.trajectory = trajectory(model, y0 = c(5, -10), tlim = c(0, 10), 
                              parameters = pars, state.names = c("x", "y"))
```

# Calulate the x and y over time with varying start conditions
```{r}
# perform numerical simulation and plot end result
model.numericalSolution = numericalSolution(model, y0 = c(-5, -5), 
                                            tlim = c(0, 20), type = "one",
                                            parameters = pars, 
                                            col = c("green", "orange"), 
                                            ylim = ylim, add.legend = F)
# make legend for plot
legend("topright", col = c("green", "orange"), legend = c("x", "y"), lty = 1)
eq = locator(n = 1)

model.numericalSolution = numericalSolution(model, y0 = c(5, 10), 
                                            tlim = c(0, 20), type = "one",
                                            parameters = pars, 
                                            col = c("green", "orange"), 
                                            ylim = ylim, add.legend = F)
# make legend for plot
legend("topright", col = c("green", "orange"), legend = c("x", "y"), lty = 1)


model.numericalSolution = numericalSolution(model, y0 = c(-5, 5), 
                                            tlim = c(0, 20), type = "one",
                                            parameters = pars, 
                                            col = c("green", "orange"), 
                                            ylim = ylim, add.legend = F)
# make legend for plot
legend("topright", col = c("green", "orange"), legend = c("x", "y"), lty = 1)


model.numericalSolution = numericalSolution(model, y0 = c(0, 5), 
                                            tlim = c(0, 20), type = "one",
                                            parameters = pars, 
                                            col = c("green", "orange"), 
                                            ylim = ylim, add.legend = F)
# make legend for plot
legend("topright", col = c("green", "orange"), legend = c("x", "y"), lty = 1)


model.numericalSolution = numericalSolution(model, y0 = c(0, -5), 
                                            tlim = c(0, 20), type = "one",
                                            parameters = pars, 
                                            col = c("green", "orange"), 
                                            ylim = ylim, add.legend = F)
# make legend for plot
legend("topright", col = c("green", "orange"), legend = c("x", "y"), lty = 1)

```




# Question 1: Calculate the stability at the selected equilibrium
```{r}

stab = stability(model, parameters = pars, system = "two.dim",
                 ystar=c(0, 0), state.names = c("x", "y"))
```









# Questions 4 - 8

```{r}
model4 = function(timepoint, state, parameters) { 
  with(as.list(c(state, parameters)), {
    dx = -x + 2*y 
    dy = -5*x - y 
    list(c(dx, dy))
  }) 
}

model5 = function(timepoint, state, parameters) { 
  with(as.list(c(state, parameters)), {
    dx = -3*x - 2*y 
    dy = -2*x - y 
    list(c(dx, dy))
  }) 
}

model6 = function(timepoint, state, parameters) { 
  with(as.list(c(state, parameters)), {
    dx = 2*x + y 
    dy = -6*x - 2*y 
    list(c(dx, dy))
  }) 
}


model7 = function(timepoint, state, parameters) { 
  with(as.list(c(state, parameters)), {
    dx = x + 3*y 
    dy = -3*x + 2*y 
    list(c(dx, dy))
  }) 
}

model8 = function(timepoint, state, parameters) { 
  with(as.list(c(state, parameters)), {
    dx = 2*x + 0.5*y 
    dy = -2*x + 5*y 
    list(c(dx, dy))
  }) 
}


# for setting the axis limits
xlim = c(-10, 10) 
ylim = c(-10, 10)

# no parameters in the ODE, so we set pars as an empty vector
pars = vector()

# It is possible to do all models in a loop, ithout changing the equilibrium, because the equiolibrium is the same for all models

for (mod in c(model4, model5, model6, model7, model8)) {
  
  # draw vector field
  model.flowField = flowField(xlab = "x", ylab = "y", mod, 
                              xlim = xlim, ylim = ylim, 
                              parameters = pars, 
                              points = 19, add = FALSE, 
                              state.names = c("x", "y"))

  # draw nullclines
  model.nullclines = nullclines(mod, xlim = xlim, ylim = ylim,
                                parameters = pars,
                                points = 500, state.names = c("x", "y"))

  # draw trajectory
  #for (start in c( c(5,5), c(-5,-5))) {
  #  model.trajectory = trajectory(mod, y0 = c(start[1], start[2]), 
  #                                tlim = c(0, 10), 
  #                                parameters = pars, state.names = c("x", "y"))
  #}
  
  model.trajectory = trajectory(mod, y0 = c(-5, -5), 
                                tlim = c(0, 10), 
                                parameters = pars, state.names = c("x", "y"))
  model.trajectory = trajectory(mod, y0 = c(-5, 5), 
                                tlim = c(0, 10), 
                                parameters = pars, state.names = c("x", "y"))
  model.trajectory = trajectory(mod, y0 = c(5, -5), 
                                tlim = c(0, 10), 
                                parameters = pars, state.names = c("x", "y"))
  model.trajectory = trajectory(mod, y0 = c(5, 5), 
                                tlim = c(0, 10), 
                                parameters = pars, state.names = c("x", "y"))
  model.trajectory = trajectory(mod, y0 = c(5, 0), 
                                tlim = c(0, 10), 
                                parameters = pars, state.names = c("x", "y"))
  model.trajectory = trajectory(mod, y0 = c(-5, 0), 
                                tlim = c(0, 10), 
                                parameters = pars, state.names = c("x", "y"))
                                
  # Calculate the stability at the selected equilibrium
  stab = stability(mod, parameters = pars, system = "two.dim",
                 ystar=c(0, 0), state.names = c("x", "y"))
  
}
```

# Question 10 NFkB oscillations with S = 0

```{r}
nfkb = function(timepoint, state, parameters) { 
  with(as.list(c(state, parameters)), {
    dI = -S + a*N - b*I 
    dN = -c*I - d*N 
    list(c(dI, dN))
  }) 
}

pars = c(a=1, b=1, c=1, d=1, S=0)

# for setting the axis limits
xlim = c(-10, 10) 
ylim = c(-10, 10)

# draw vector field
model.flowField = flowField(xlab = "I", ylab = "N", nfkb, 
                            xlim = xlim, ylim = ylim, 
                            parameters = pars, 
                            points = 19, add = FALSE, 
                            state.names = c("I", "N"))

# draw nullclines
model.nullclines = nullclines(nfkb, xlim = xlim, ylim = ylim,
                              parameters = pars,
                              points = 500, state.names = c("I", "N"))

# draw trajectory
model.trajectory = trajectory(nfkb, y0 = c(-5, -5), tlim = c(0, 10), 
                              parameters = pars, state.names = c("I", "N"))

# Calulate the N and I over time
model.numericalSolution = numericalSolution(nfkb, y0 = c(-5, -5), 
                                            tlim = c(0, 20), type = "one",
                                            parameters = pars, 
                                            state.names = c("I", "N"),
                                            col = c("green", "orange"), 
                                            ylim = ylim, add.legend = F)
# make legend for plot
legend("topright", col = c("green", "orange"), legend = c("x", "y"), lty = 1)

# Determine stability
stab = stability(nfkb, parameters = pars, system = "two.dim",
                 ystar=c(0, 0), state.names = c("I", "N"))

```

# Question 11 NFkB oscillations with S != 0

```{r}
nfkb = function(timepoint, state, parameters) { 
  with(as.list(c(state, parameters)), {
    dI = -S + a*N - b*I 
    dN = -c*I - d*N 
    list(c(dI, dN))
  }) 
}

pars = c(a=1, b=1, c=1, d=1, S=10)

# for setting the axis limits
xlim = c(-10, 10) 
ylim = c(-10, 10)

# draw vector field
model.flowField = flowField(xlab = "I", ylab = "N", nfkb, 
                            xlim = xlim, ylim = ylim, 
                            parameters = pars, 
                            points = 19, add = FALSE, 
                            state.names = c("I", "N"))

# draw nullclines
model.nullclines = nullclines(nfkb, xlim = xlim, ylim = ylim,
                              parameters = pars,
                              points = 500, state.names = c("I", "N"))

# draw trajectory
model.trajectory = trajectory(nfkb, y0 = c(-5, -5), tlim = c(0, 10), 
                              parameters = pars, state.names = c("I", "N"))

# Calulate the N and I over time
model.numericalSolution = numericalSolution(nfkb, y0 = c(-5, -5), 
                                            tlim = c(0, 20), type = "one",
                                            parameters = pars, 
                                            state.names = c("I", "N"),
                                            col = c("green", "orange"), 
                                            ylim = ylim, add.legend = F)
# make legend for plot
legend("topright", col = c("green", "orange"), legend = c("x", "y"), lty = 1)

# Determine stability
stab = stability(nfkb, parameters = pars, system = "two.dim",
                 ystar=c(0, 0), state.names = c("I", "N"))

```

# Question 16 Auto inhibition for NFkB and IkB aligned and change together

Not clear how to model this best

```{r}
pars = c(a=3, b=1, c=2, d=1, S=0)

# for setting the axis limits
xlim = c(-10, 10) 
ylim = c(-40, 40)

# draw vector field
model.flowField = flowField(xlab = "I", ylab = "N", nfkb, 
                            xlim = xlim, ylim = ylim, 
                            parameters = pars, 
                            points = 19, add = FALSE, 
                            state.names = c("I", "N"))

# draw nullclines
model.nullclines = nullclines(nfkb, xlim = xlim, ylim = ylim,
                              parameters = pars,
                              points = 500, state.names = c("I", "N"))

# draw trajectory
model.trajectory = trajectory(nfkb, y0 = c(-5, -5), tlim = c(0, 10), 
                              parameters = pars, state.names = c("I", "N"))

# Calulate the N and I over time
model.numericalSolution = numericalSolution(nfkb, y0 = c(-5, -5), 
                                            tlim = c(0, 20), type = "one",
                                            parameters = pars, 
                                            state.names = c("I", "N"),
                                            col = c("green", "orange"), 
                                            ylim = ylim, add.legend = F)
# make legend for plot
legend("topright", col = c("green", "orange"), legend = c("x", "y"), lty = 1)

# Determine stability
stab = stability(nfkb, parameters = pars, system = "two.dim",
                 ystar=c(0, 0), state.names = c("I", "N"))
```


# Question 17 Auto inhibition removed

This means that parameters b and d become 0

```{r}
pars = c(a=1, b=0, c=1, d=0, S=10)

# for setting the axis limits
xlim = c(-10, 10) 
ylim = c(-40, 40)

# draw vector field
model.flowField = flowField(xlab = "I", ylab = "N", nfkb, 
                            xlim = xlim, ylim = ylim, 
                            parameters = pars, 
                            points = 19, add = FALSE, 
                            state.names = c("I", "N"))

# draw nullclines
model.nullclines = nullclines(nfkb, xlim = xlim, ylim = ylim,
                              parameters = pars,
                              points = 500, state.names = c("I", "N"))

# draw trajectory
model.trajectory = trajectory(nfkb, y0 = c(-5, -5), tlim = c(0, 10), 
                              parameters = pars, state.names = c("I", "N"))

# Calulate the N and I over time
model.numericalSolution = numericalSolution(nfkb, y0 = c(-5, -5), 
                                            tlim = c(0, 20), type = "one",
                                            parameters = pars, 
                                            state.names = c("I", "N"),
                                            col = c("green", "orange"), 
                                            ylim = ylim, add.legend = F)
# make legend for plot
legend("topright", col = c("green", "orange"), legend = c("x", "y"), lty = 1)

# Determine stability
stab = stability(nfkb, parameters = pars, system = "two.dim",
                 ystar=c(0, 0), state.names = c("I", "N"))
```


# Question 18 Negative Auto inhibition 

This means that parameters b and d become negative. You get immediate oscillating instability

```{r}
pars = c(a=1, b=-1, c=1, d=-1, S=10)

# for setting the axis limits
xlim = c(-10, 10) 
ylim = c(-50, 50)

# draw vector field
model.flowField = flowField(xlab = "I", ylab = "N", nfkb, 
                            xlim = xlim, ylim = ylim, 
                            parameters = pars, 
                            points = 19, add = FALSE, 
                            state.names = c("I", "N"))

# draw nullclines
model.nullclines = nullclines(nfkb, xlim = xlim, ylim = ylim,
                              parameters = pars,
                              points = 500, state.names = c("I", "N"))

# draw trajectory
model.trajectory = trajectory(nfkb, y0 = c(-5, -5), tlim = c(0, 10), 
                              parameters = pars, state.names = c("I", "N"))

# Calulate the N and I over time
ylim = c(-1000, 1000)
model.numericalSolution = numericalSolution(nfkb, y0 = c(-5, -5), 
                                            tlim = c(0, 20), type = "one",
                                            parameters = pars, 
                                            state.names = c("I", "N"),
                                            col = c("green", "orange"), 
                                            ylim = ylim, add.legend = F)
# make legend for plot
legend("topright", col = c("green", "orange"), legend = c("x", "y"), lty = 1)
```
