---
title: "CBR Assignment 5"
author: "Hans"
date: "9/27/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pacman)
p_load(tidyverse)
p_load(ggplot2)
p_load(Deriv)
p_load(deSolve)
p_load(phaseR)
p_load(gridExtra)

theme_set(theme_minimal())
```

# Question 6: 

Simulate a time plot of the model setting Vm = 1.0, K = 0.5 and the degradation parameter you chose to 0.1 and using the ode function from the R package deSolve. Set P1 at a value of 0.3. Note that hereafter we will consider pulses of P1; in order to plot a time course of P1 it is handy to now already include it as the ODE dP 1/dt = 0 and initialize it at 0.3. What happens in the long run?

```{r define_mod 23AND}
model23AND = function(timepoint, state, parameters) { 
  with(as.list(c(state, parameters)), {
    dP2 = (Vm * P1^4 ) / (K + P1^4 ) - kdeg * P2
    dP3 = (Vm * P1^4 * P2^4) / (K + P1^4 * P2^4) - kdeg * P3 
    list(c(dP2, dP3))
  }) 
}
```

```{r define_model 123AND}
model123AND = function(timepoint, state, parameters) { 
  with(as.list(c(state, parameters)), {
    dP1 = 0
    dP2 = (Vm * P1^4 ) / (K + P1^4 ) - kdeg * P2
    dP3 = (Vm * P1^4 * P2^4) / (K + P1^4 * P2^4) - kdeg * P3 
    list(c(dP1, dP2, dP3))
  }) 
}
```

```{r }
pars = c(Vm = 1.0, K = 0.5, P1 = 0.3, kdeg = 0.1)
inistate = c(P1 = 0.3, P2 = 0, P3 = 0)

time_int = c(0, 100)

# Set the axis limits
xlim = c(-10, 10) 
ylim = c(-10, 10)

model.flowField = flowField(model23AND, 
                            xlim = xlim, 
                            ylim = ylim, 
                            parameters = pars,
                            state.names = c("P2", "P3"),
                            points = 19, 
                            add = FALSE, 
                            xlab = "P2", 
                            ylab = "P3")

# draw nullclines
model.nullclines = nullclines(model23AND, 
                              xlim = xlim, 
                              ylim = ylim,
                              parameters = pars,
                              points = 500, 
                              state.names = c("P2", "P3"))

# draw trajectory
inival = c(5, 5)
model.trajectory = trajectory(model23AND, 
                              y0 = inival, 
                              tlim = time_int, 
                              parameters = pars, 
                              state.names = c("P2", "P3"))
inival = c(0, 5)
model.trajectory = trajectory(model23AND, 
                              y0 = inival, 
                              tlim = time_int, 
                              parameters = pars, 
                              state.names = c("P2", "P3"))
inival = c(5, 0)
model.trajectory = trajectory(model23AND, 
                              y0 = inival, 
                              tlim = time_int, 
                              parameters = pars, 
                              state.names = c("P2", "P3"))

```

```{r}

inistate = c(P1 = 0.3, P2 = 0, P3 = 0)
time_series = seq(0, 100, by = 0.1)
pars = c(Vm = 1.0, K = 0.5, kdeg = 0.1)

out = ode(y = inistate, 
          times = time_series, 
          func = model123AND, 
          parms = pars)

df = data.frame(out)

p1 <- ggplot(data = df) +
  geom_line(mapping = aes(x=time, y = P1)) 
 
p2 <- ggplot(data = df) +
  geom_line(mapping = aes(x=time, y = P2))
  
p3 <- ggplot(data = df) +
  geom_line(mapping = aes(x=time, y = P3)) 

gridExtra::grid.arrange(
    p1,
    p2,
    p3,
    nrow = 2)

```

# Question 7: 

Next we consider the cell to receive either a brief or a somewhat longer stimulus reaching this part of the network, which enters the network via activation of P1. We model this by setting P1 briefly to a higher level, after which it instantaneously returns to its baseline level. Specifically, for the brief pulse: from time 0-10 min, P1 is at level 0.3 as before, from 10-11 min it is at level 0.7, and from 11-40 min it is at level 0.3 again. For the long pulse, the level of 0.7 is maintained from 10-15 min. Simulate both cases by splitting in three parts and combining the outputs to make a single time course plot.

```{r Simulation of short puls AND model}

pars = c(Vm = 1.0, K = 0.5, kdeg = 0.1)

time_series1 = seq(0,  10, by = 0.1)
inistate = c(P1 = 0.3, P2 = 0, P3 = 0)
out1 = ode(y = inistate, 
           times = time_series1, 
           func = model123AND, 
           parms = pars)
P2_end = out1[nrow(out1),"P2"]
P3_end = out1[nrow(out1),"P3"]

time_series2 = seq(10, 11, by = 0.1)  
inistate = c(P1 = 0.7, P2 = as.numeric(P2_end), P3 = as.numeric(P3_end))
out2 = ode(y = inistate, 
           times = time_series2, 
           func = model123AND, 
           parms = pars)

P2_end = as.numeric(out2[nrow(out2),"P2"])
P3_end = as.numeric(out2[nrow(out2),"P3"])

time_series3 = seq(11, 40, by = 0.1)
inistate = c(P1 = 0.3, P2 = P2_end, P3 = P3_end)
out3 = ode(y = inistate, 
           times = time_series3, 
           func = model123AND, 
           parms = pars)

out <- rbind(out1, out2, out3)

p1 <- ggplot(data = data.frame(out)) +
  geom_line(mapping = aes(x=time, y = P1)) 
 
p2 <- ggplot(data = data.frame(out)) +
  geom_line(mapping = aes(x=time, y = P2))
  
p3 <- ggplot(data = data.frame(out)) +
  geom_line(mapping = aes(x=time, y = P3)) 

gridExtra::grid.arrange(
    p1,
    p2,
    p3,
    nrow = 2)
```

```{r Simulation of long puls AND model}

pars = c(Vm = 1.0, K = 0.5, kdeg = 0.1)

time_series1 = seq(0,  10, by = 0.1)
inistate = c(P1 = 0.3, P2 = 0, P3 = 0)
out1 = ode(y = inistate, 
           times = time_series1, 
           func = model123AND, 
           parms = pars)
P2_end = out1[nrow(out1),"P2"]
P3_end = out1[nrow(out1),"P3"]

time_series2 = seq(10, 15, by = 0.1)  
inistate = c(P1 = 0.7, P2 = as.numeric(P2_end), P3 = as.numeric(P3_end))
out2 = ode(y = inistate, 
           times = time_series2, 
           func = model123AND, 
           parms = pars)

P2_end = as.numeric(out2[nrow(out2),"P2"])
P3_end = as.numeric(out2[nrow(out2),"P3"])

time_series3 = seq(15, 40, by = 0.1)
inistate = c(P1 = 0.3, P2 = P2_end, P3 = P3_end)
out3 = ode(y = inistate, 
           times = time_series3, 
           func = model123AND, 
           parms = pars)

out <- rbind(out1, out2, out3)

p1 <- ggplot(data = data.frame(out)) +
  geom_line(mapping = aes(x=time, y = P1)) 
 
p2 <- ggplot(data = data.frame(out)) +
  geom_line(mapping = aes(x=time, y = P2))
  
p3 <- ggplot(data = data.frame(out)) +
  geom_line(mapping = aes(x=time, y = P3)) 

gridExtra::grid.arrange(
    p1,
    p2,
    p3,
    nrow = 2)

```


# Question 13  OR Model

Again simulate time plots for 1-min and 5-min pulses.

```{r First the simulation without pulse}

model123OR = function(timepoint, state, parameters) { 
  with(as.list(c(state, parameters)), {
    dP1 = 0
    dP2 = (Vm * P1^4 ) / (K + P1^4 ) - kdeg * P2 
    dP3 = (Vm * (P1^4 + P2^4)) / (K + (P1^4 + P2^4)) - kdeg * P3 
    list(c(dP1, dP2, dP3))
  }) 
}

inistate = c(P1 = 0.3, P2 = 0, P3 = 0)
time_series = seq(0, 100, by = 0.1)
pars = c(Vm = 1.0, K = 0.5, kdeg = 0.1)

out = ode(y = inistate, 
          times = time_series, 
          func = model123OR, 
          parms = pars)

df = data.frame(out)

p1 <- ggplot(data = df) +
  geom_line(mapping = aes(x=time, y = P1)) 
 
p2 <- ggplot(data = df) +
  geom_line(mapping = aes(x=time, y = P2))
  
p3 <- ggplot(data = df) +
  geom_line(mapping = aes(x=time, y = P3)) 

gridExtra::grid.arrange(
    p1,
    p2,
    p3,
    nrow = 2)
```

```{r Simulation of short pulsOR model}

pars = c(Vm = 1.0, K = 0.5, kdeg = 0.1)

time_series1 = seq(0,  10, by = 0.1)
inistate = c(P1 = 0.3, P2 = 0, P3 = 0)
out1 = ode(y = inistate, 
           times = time_series1, 
           func = model123OR, 
           parms = pars)
P2_end = out1[nrow(out1),"P2"]
P3_end = out1[nrow(out1),"P3"]

time_series2 = seq(10, 11, by = 0.1)  
inistate = c(P1 = 0.7, P2 = as.numeric(P2_end), P3 = as.numeric(P3_end))
out2 = ode(y = inistate, 
           times = time_series2, 
           func = model123OR, 
           parms = pars)

P2_end = as.numeric(out2[nrow(out2),"P2"])
P3_end = as.numeric(out2[nrow(out2),"P3"])

time_series3 = seq(11, 40, by = 0.1)
inistate = c(P1 = 0.3, P2 = P2_end, P3 = P3_end)
out3 = ode(y = inistate, 
           times = time_series3, 
           func = model123OR, 
           parms = pars)

out <- rbind(out1, out2, out3)

p1 <- ggplot(data = data.frame(out)) +
  geom_line(mapping = aes(x=time, y = P1)) 
 
p2 <- ggplot(data = data.frame(out)) +
  geom_line(mapping = aes(x=time, y = P2))
  
p3 <- ggplot(data = data.frame(out)) +
  geom_line(mapping = aes(x=time, y = P3)) 

gridExtra::grid.arrange(
    p1,
    p2,
    p3,
    nrow = 2)
```

```{r Simulation of long puls OR Model}

pars = c(Vm = 1.0, K = 0.5, kdeg = 0.1)

time_series1 = seq(0,  10, by = 0.1)
inistate = c(P1 = 0.3, P2 = 0, P3 = 0)
out1 = ode(y = inistate, 
           times = time_series1, 
           func = model123OR, 
           parms = pars)
P2_end = out1[nrow(out1),"P2"]
P3_end = out1[nrow(out1),"P3"]

time_series2 = seq(10, 15, by = 0.1)  
inistate = c(P1 = 0.7, P2 = as.numeric(P2_end), P3 = as.numeric(P3_end))
out2 = ode(y = inistate, 
           times = time_series2, 
           func = model123OR, 
           parms = pars)

P2_end = as.numeric(out2[nrow(out2),"P2"])
P3_end = as.numeric(out2[nrow(out2),"P3"])

time_series3 = seq(15, 40, by = 0.1)
inistate = c(P1 = 0.3, P2 = P2_end, P3 = P3_end)
out3 = ode(y = inistate, 
           times = time_series3, 
           func = model123OR, 
           parms = pars)

out <- rbind(out1, out2, out3)

p1 <- ggplot(data = data.frame(out)) +
  geom_line(mapping = aes(x=time, y = P1)) 
 
p2 <- ggplot(data = data.frame(out)) +
  geom_line(mapping = aes(x=time, y = P2))
  
p3 <- ggplot(data = data.frame(out)) +
  geom_line(mapping = aes(x=time, y = P3)) 

gridExtra::grid.arrange(
    p1,
    p2,
    p3,
    nrow = 2)

```


