---
title: "CBR assignment 3"
author: "Hans"
date: "9/19/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pacman)
p_load(tidyverse)
p_load(ggplot2)
p_load(GillespieSSA)
p_load(deSolve)
p_load(phaseR)
theme_set(theme_minimal())
source("killingfunctions.R")
```



# Question 2

```{r Question 2}

InitialCellCount <- 25

# For the one_hit_ODE you need a tinestep of 0.5 s to see the individual cell deaths 
parms <- c <- c(k1 = 0.8, k2 = 0.8, k3 = 0.8)
times <- seq(0, 12, 0.5)
out.ode <- one_hit_ODE(times, parms, InitialCellCount)
plot(out.ode)

# For the one_hit_ODE with a tinestep of 0.1 s you get a connected line 
parms <- c <- c(k1 = 0.8, k2 = 0.8, k3 = 0.8)
times <- seq(0, 12, 0.1)
out.ode <- one_hit_ODE(times, parms, InitialCellCount)
plot(out.ode)

# For Gillespie the time step does does not make a difference: time is b=not a parameter
parms <- c <- c(k1 = 0.8, k2 = 0.8, k3 = 0.8)
out.gil <- one_hit_gillespie(parms, InitialCellCount)
plot(out.gil)
```

# Question 3: 

Which model do you think is the best choice to represent the discussed experimental data? Why?

Solution: The Gillespie simulations because at such low numbers of target cells the deaths are still visible as random events and we are specifically interested in the variability in number of deaths across the wells.

# Question 4

```{r Question 4}

InitialCellCount <- 25
simtype <- "one_hit_ODE"
times <- seq(0, 12, 0.1)
nwells <- 150
stdev <- 2

# Its is a matter of trying paramaters and try to emulate the pattern....

InitialCellCount <- 25
parms <- c(k1 = 1.5, k2 = 3, k3 = 0.5)
out <- multiWell(nwells, simtype, parms, InitialCellCount, times, stdev )
wellPlot(out)
barchartplotter(out)

InitialCellCount <- 10
parms <- c(k1 = 1.5, k2 = 3, k3 = 0.5)
out <- multiWell(nwells, simtype, parms, InitialCellCount, times, stdev )
wellPlot(out)
barchartplotter(out)

```


# Question 5: 

Now use the multiWell function to simulate 150 wells with the Gillespie simulation. Take k1 = 1.5, k2 = 3 and vary k3 and the initial amount of targets until the distribution of deaths roughly matches that of the Poisson distribution above. Plot the outputs as bar charts and as scatterplots. How well do these simulated data match the experimental data? Which characteristics do match and which do not?

```{r Question 5}

simtype <- "one_hit_gillespie"
times <- seq(0, 12, 0.1)
nwells <- 150
stdev <- 2

# Its is a matter of varying InitialCellCount and k3  and try to emulate the pattern....

InitialCellCount <- 25
parms <- c <- c(k1 = 1.5, k2 = 3, k3 = 0.8)
out <- multiWell(nwells, simtype, parms, InitialCellCount, times, stdev )
wellPlot(out)
barchartplotter(out)

InitialCellCount <- 10
parms <- c <- c(k1 = 1.5, k2 = 3, k3 = 0.8)
out <- multiWell(nwells, simtype, parms, InitialCellCount, times, stdev )
wellPlot(out)
barchartplotter(out)

InitialCellCount <- 10
parms <- c <- c(k1 = 1.5, k2 = 3, k3 = 0.5)
out <- multiWell(nwells, simtype, parms, InitialCellCount, times, stdev )
wellPlot(out)
barchartplotter(out)
```


# Question 7 

You can run a two hit ODE model by using the simtype "two_hit_ODE" or a two hit Gillespie simulation by using the simtype "two_hit_gillespie". Use the models described above to get as close to the experimental data as possible. When you have decided on a model and parameters, run that simulation a few times to characterise how much variation there is in the results. Can you get closer to the experimental results than with the one-hit model?

```{r Question 7}


times <- seq(0, 12, 0.1)
nwells <- 150
stdev <- 2

# Its is a matter of varying InitialCellCount and k3  and try to emulate the pattern....

# First ODE
simtype <- "two_hit_ODE"

InitialCellCount <- 25
stdev <- 5
parms <- c(k1 = 1.5, k2 = 3, k3 = 0.8)
out <- multiWell(nwells, simtype, parms, InitialCellCount, times, stdev )
wellPlot(out)
barchartplotter(out)

InitialCellCount <- 10
stdev <- 5
parms <- c (k1 = 1.5, k2 = 3, k3 = 0.8)
out <- multiWell(nwells, simtype, parms, InitialCellCount, times, stdev )
wellPlot(out)
barchartplotter(out)

InitialCellCount <- 25
stdev <- 5
parms <- c(k1 = 0.5, k2 = 0.5, k3 = 0.2)
out <- multiWell(nwells, simtype, parms, InitialCellCount, times, stdev )
wellPlot(out)
barchartplotter(out)


# Then Gillespie

simtype <- "two_hit_gillespie"

InitialCellCount <- 25
stdev <- 5
parms <- c(k1 = 1.5, k2 = 3, k3 = 0.8)
out <- multiWell(nwells, simtype, parms, InitialCellCount, times, stdev )
wellPlot(out)
barchartplotter(out)

InitialCellCount <- 10
stdev <- 5
parms <- c(k1 = 1.5, k2 = 3, k3 = 0.8)
out <- multiWell(nwells, simtype, parms, InitialCellCount, times, stdev )
wellPlot(out)
barchartplotter(out)

InitialCellCount <- 10
stdev <- 5
parms <- c(k1 = 0.5, k2 = 0.5, k3 = 0.2)
out <- multiWell(nwells, simtype, parms, InitialCellCount, times, stdev )
wellPlot(out)
barchartplotter(out)
```

# Question 10

```{r Question 10}
source("csc.r")
parameters = c(p1 = 0.5, p2 = 0.1, 
               q1 = 0.5, q2 = 0.1, 
               rcsc = 1, rpc = 1, 
               dcsc = 0, dpc = 0, ddc = 0.5)
```


# Question 12

Create a vector with the inital amount of cells based on the above-defined CSC fraction and distribute the remaining cells evenly over PCs and DCs.

```{r Question 12}
# number of cells at initialization
ncells = 1000

# cancer stem cell fraction
csc0 = 0.1

yinit = c(CSC = csc0 * ncells, 
          PC  = 0.5 * (1 - csc0) * ncells, 
          DC  = 0.5 * (1 - csc0) * ncells)
```



# Question 13

Create a time sequence and run the model using the predefined function run_model which expects three arguments: a time sequence, a vector for the initial condition and one for the parameters. Make sure you are able to change the model parameters and initial condition. What happens during the simulation? Does it settle at an equilibrium?

```{r Question 13}

times = seq(0, 50, by = 0.01) 

run_model(times, yinit, parameters)


```

# Question 14

Test how the initial stem cell fraction affects tumour growth. Consider both the absolute and relative number of CSCs, PCs, and DCs during tumour growth.

You find that the precentage CSC does not change much as wuth changinbg cancer stem cell fraction (and therefire woith increasing number of initial stem cells)

```{r}
# cancer stem cell fraction
csc0 = 0.1
yinit = c(CSC = csc0 * ncells, 
          PC  = 0.5 * (1 - csc0) * ncells, 
          DC  = 0.5 * (1 - csc0) * ncells)
run_model(times, yinit, parameters)

csc0 = 0.25
yinit = c(CSC = csc0 * ncells, 
          PC  = 0.5 * (1 - csc0) * ncells, 
          DC  = 0.5 * (1 - csc0) * ncells)
run_model(times, yinit, parameters)

csc0 = 0.5
yinit = c(CSC = csc0 * ncells, 
          PC  = 0.5 * (1 - csc0) * ncells, 
          DC  = 0.5 * (1 - csc0) * ncells)
run_model(times, yinit, parameters)

csc0 = 0.75
yinit = c(CSC = csc0 * ncells, 
          PC  = 0.5 * (1 - csc0) * ncells, 
          DC  = 0.5 * (1 - csc0) * ncells)
run_model(times, yinit, parameters)

csc0 = 0.01
yinit = c(CSC = csc0 * ncells, 
          PC  = 0.5 * (1 - csc0) * ncells, 
          DC  = 0.5 * (1 - csc0) * ncells)
run_model(times, yinit, parameters)

```



# Question 16 + 17 + 18

Use the derived inequality to run the model with parameters such that the CSC population grows, shrinks, or remains constant. If the CSC population does not develop as expected, re-evaluate the inequality.

Refer to Word file

(2p1 - 1 - p2) * rcsc - dcsc  = 0   -> Constant
(2p1 - 1 - p2) * rcsc - dcsc  > 0   -> Growth
(2p1 - 1 - p2) * rcsc - dcsc  < 0   -> Decline

```{r Question 18}

times = seq(0, 50, by = 0.01) 

# An example for a declining CSC population:
parameters = c(p1 = 0.5, p2 = 0.1, 
               q1 = 0.5, q2 = 0.1, 
               rcsc = 1, rpc = 1, 
               dcsc = 0.2, dpc = 0, ddc = 0.5)

((2 * parameters['p1'] - 1 + parameters['p2']) * parameters['rcsc'] - parameters['dcsc'])

run_model(times, yinit, parameters)

# An example for a constant CSC population:
parameters = c(p1 = 0.5, p2 = 0.1, 
               q1 = 0.5, q2 = 0.1, 
               rcsc = 1, rpc = 1, 
               dcsc = 0.1, dpc = 0, ddc = 0.5)
((2 * parameters['p1'] - 1 + parameters['p2']) * parameters['rcsc'] - parameters['dcsc'])
run_model(times, yinit, parameters)

# An example for an increasing  CSC population:
parameters = c(p1 = 0.5, p2 = 0.1, 
               q1 = 0.5, q2 = 0.1, 
               rcsc = 1, rpc = 1, 
               dcsc = 0, dpc = 0, ddc = 0.5)
((2 * parameters['p1'] - 1 + parameters['p2']) * parameters['rcsc'] - parameters['dcsc'])
run_model(times, yinit, parameters)
```


# Question 19 + 20 + 21

Use the derived inequality to run the model with parameters such that the PC population grows, shrinks, or remains constant (still in the absence of CSCs). If the PC population does not develop as expected, re-evaluate the inequality.

Refer to Word file

(2q1 - 1 + q2) * rpc - dpc  = 0   -> Constant
(2q1 - 1 + q2) * rpc - dpc  > 0   -> Growth
(2q1 - 1 + q2) * rpc - dpc  < 0   -> Decline

```{r Question 21}

times = seq(0, 50, by = 0.01) 
csc0 = 0
yinit = c(CSC = csc0 * ncells, 
          PC = 0.5 * (1 - csc0) * ncells, 
          DC = 0.5 * (1 - csc0) * ncells)

# An example for a declining PC population:

parameters = c(p1 = 0.5, p2 = 0.1, 
               q1 = 0.5, q2 = 0.1, 
               rcsc = 1, rpc = 1, 
               dcsc = 0.2, dpc = 0.2, ddc = 0.5)
((2 * parameters['q1'] - 1 + parameters['q2']) * parameters['rpc'] - parameters['dpc'])
run_model(times, yinit, parameters)

# An example for a constant PC population:
parameters = c(p1 = 0.5, p2 = 0.1, 
               q1 = 0.5, q2 = 0.1, 
               rcsc = 1, rpc = 1, 
               dcsc = 0.2, dpc = 0.1, ddc = 0.5)
((2 * parameters['q1'] - 1 + parameters['q2']) * parameters['rpc'] - parameters['dpc'])
run_model(times, yinit, parameters)

# An example for an increasing  PC population:
parameters = c(p1 = 0.5, p2 = 0.1, 
               q1 = 0.5, q2 = 0.2, 
               rcsc = 1, rpc = 1, 
               dcsc = 0, dpc = 0, ddc = 0.5)
((2 * parameters['q1'] - 1 + parameters['q2']) * parameters['rpc'] - parameters['dpc'])
run_model(times, yinit, parameters)
```

# Question 22

You derived two conditions, one determining whether the number of CSCs increases and one determining whether the number of PCs increases. Investigate what happens to the total size of the tumour population when the parameters do not allow for growth of the CSC and PC population. Why does this happen?


```{r Question 22}
times = seq(0, 500, by = 0.1) 
csc0 = 0.1
ncells = 1000
yinit = c(CSC = csc0 * ncells, 
          PC = 0.5 * (1 - csc0) * ncells, 
          DC = 0.5 * (1 - csc0) * ncells)

# An example for a declining PC population:

parameters = c(p1 = 0.5, p2 = 0.1, 
               q1 = 0.5, q2 = 0.1, 
               rcsc = 1, rpc = 1, 
               dcsc = 0.2, dpc = 0.2, ddc = 0.5)

((2 * parameters['p1'] - 1 + parameters['p2']) * parameters['rcsc'] - parameters['dcsc'])
((2 * parameters['q1'] - 1 + parameters['q2']) * parameters['rpc'] - parameters['dpc'])

run_model(times, yinit, parameters)
```


# Question 23

You found conditions for which the population either grows or shrinks, there are however also parameters for which the population reaches an equilibrium size with all types of cells present. For this, we will search for parameters for which dCSC = dPC = dDC = 0. To simplify this search, we assume that CSCs and PCs do not die, all cells divide at rate 1, and there are no asymmetric stem cell divisions. Solve the dCSC equation to find values for p1 and p3 such that the CSC population is at equilibrium.

p1 = p3 and p2 = 0, so p1 = p3 = 0.5

# Question 24
What is the expression for the equilibrium value for CSC, i.e., what is the value for CSC if dCSC = 0

Solution: The solution of dCSC = 0 is independent of CSC. Hence, CSC never changes and is thus equal to the initial value, which we call CSC0.

# Question 27
Find values for q1, q2, and q3 such that the equilibrium value for PC is positive and substitute those values in the result of the previous question. Use this result to predict the effect of dDC on the DC equilibrium value and test this by simulating the model.


```{r}
csc0 = 0.1
CSC0 = csc0 * ncells
yinit = c(CSC = csc0 * ncells, 
          PC = 0.5 * (1 - csc0) * ncells, 
          DC = 0.5 * (1 - csc0) * ncells)
parameters = c(p1 = 0.5, p2 = 0, 
               q1 = 0.4, q2 = 0.1, 
               rcsc = 1, rpc = 1, 
               dcsc = 0, dpc = 0, ddc = 0.3)
times = seq(0, 100, by = 0.1) 
run_model(times, yinit, parameters)

q3 = 1 - parameters['q1'] - parameters['q2']
(PC <- CSC0 / (q3 - parameters['q1']))
(DC <- (parameters['q2'] + 2 * q3) * PC /  parameters['ddc'])
(CSC0)
```

