---
title: "Assignment 1"
author: "Hans"
date: "9/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("phaseR")
library(tidyverse)
```

Single analysis
phasePlaneAnalysis(model, xlim = xlim, ylim = ylim, parameters = pars)

# Define the model and plot flow field, null clines and one trajectory
```{r define_model}


model = function(timepoint, state, parameters) { 
  with(as.list(c(state, parameters)), {
    dx = -2 * x - 3 * y 
    dy = -x - 2 * y 
    list(c(dx, dy))
  }) 
}

# for setting the axis limits
xlim = c(-10, 10) 
ylim = c(-10, 10)

# no parameters in the ODE, so we set pars as an empty vector
pars = vector()
# draw vector field
model.flowField = flowField(xlab = "x", 
                            ylab = "y", 
                            model, 
                            xlim = xlim, 
                            ylim = ylim, 
                            parameters = pars, 
                            points = 19, 
                            add = FALSE, 
                            state.names = c("x", "y"))

# draw nullclines
model.nullclines = nullclines(model, xlim = xlim, ylim = ylim,
                              parameters = pars,
                              points = 500, state.names = c("x", "y"))

# draw trajectory
model.trajectory = trajectory(model, y0 = c(-5, -5), tlim = c(0, 10), 
                              parameters = pars, state.names = c("x", "y"))
model.trajectory = trajectory(model, y0 = c(0, -10), tlim = c(0, 10), 
                              parameters = pars, state.names = c("x", "y"))
model.trajectory = trajectory(model, y0 = c(-5, 5), tlim = c(0, 10), 
                              parameters = pars, state.names = c("x", "y"))
model.trajectory = trajectory(model, y0 = c(5, -10), tlim = c(0, 10), 
                              parameters = pars, state.names = c("x", "y"))
```

# Calulate the x and y over time with varying start conditions
```{r}

# Perform numerical simulation and plot end result
# The function numericalSolution

model.numericalSolution = numericalSolution(model, y0 = c(-5, -5), 
                                            tlim = c(0, 20), 
                                            type = "one",
                                            parameters = pars, 
                                            col = c("green", "orange"), 
                                            ylim = ylim, 
                                            add.legend = F)
legend("topright", col = c("green", "orange"), legend = c("x", "y"), lty = 1)
eq = locator(n = 1)

model.numericalSolution = numericalSolution(model, y0 = c(5, 10), 
                                            tlim = c(0, 20), 
                                            type = "one",
                                            parameters = pars, 
                                            col = c("green", "orange"), 
                                            ylim = ylim, 
                                            add.legend = F)
legend("topright", col = c("green", "orange"), legend = c("x", "y"), lty = 1)


model.numericalSolution = numericalSolution(model, y0 = c(-5, 5), 
                                            tlim = c(0, 20), 
                                            type = "one",
                                            parameters = pars, 
                                            col = c("green", "orange"), 
                                            ylim = ylim, 
                                            add.legend = F)
legend("topright", col = c("green", "orange"), legend = c("x", "y"), lty = 1)


model.numericalSolution = numericalSolution(model, y0 = c(0, 5), 
                                            tlim = c(0, 20), type = "one",
                                            parameters = pars, 
                                            col = c("green", "orange"), 
                                            ylim = ylim, 
                                            add.legend = F)
legend("topright", col = c("green", "orange"), legend = c("x", "y"), lty = 1)


model.numericalSolution = numericalSolution(model, y0 = c(0, -5), 
                                            tlim = c(0, 20), type = "one",
                                            parameters = pars, 
                                            col = c("green", "orange"), 
                                            ylim = ylim, 
                                            add.legend = F)
legend("topright", col = c("green", "orange"), legend = c("x", "y"), lty = 1)


```

```{r}
```





# Question 1: Calculate the stability at the selected equilibrium
```{r}

# ystar is on indication of where the equilibrium is
# function can be used both for one and two equations
eq = c(0, 0)
stab = stability(model, parameters = pars, system = "two.dim",
                 ystar = eq, state.names = c("x", "y"))
```



# Question 3

```{r}
# You need to do this from the console
# phasePlaneAnalysis(model, xlim = xlim, ylim = ylim, parameters = pars) 
```






# Questions 4 - 8

```{r}
model4 = function(timepoint, state, parameters) { 
  with(as.list(c(state, parameters)), {
    dx = -x + 2*y 
    dy = -5*x - y 
    list(c(dx, dy))
  }) 
}

model5 = function(timepoint, state, parameters) { 
  with(as.list(c(state, parameters)), {
    dx = -3*x - 2*y 
    dy = -2*x - y 
    list(c(dx, dy))
  }) 
}

model6 = function(timepoint, state, parameters) { 
  with(as.list(c(state, parameters)), {
    dx = 2*x + y 
    dy = -6*x - 2*y 
    list(c(dx, dy))
  }) 
}

model7 = function(timepoint, state, parameters) { 
  with(as.list(c(state, parameters)), {
    dx = x + 3*y 
    dy = -3*x + 2*y 
    list(c(dx, dy))
  }) 
}

model8 = function(timepoint, state, parameters) { 
  with(as.list(c(state, parameters)), {
    dx = 2*x + 0.5*y 
    dy = -2*x + 5*y 
    list(c(dx, dy))
  }) 
}


# for setting the axis limits
xlim = c(-10, 10) 
ylim = c(-10, 10)

# no parameters in the ODE, so we set pars as an empty vector
pars = vector()

# It is possible to do all models in a loop, ithout changing the equilibrium, because the equiolibrium is the same for all models

for (mod in c(model4, model5, model6, model7, model8)) {
  
  # draw vector field
  model.flowField = flowField(mod,
                              xlim = xlim, ylim = ylim, 
                              parameters = pars, 
                              state.names = c("x", "y"),
                              points = 19, add = FALSE, 
                              xlab = "x", ylab = "y")

  # draw nullclines
  model.nullclines = nullclines(mod, xlim = xlim, ylim = ylim,
                                parameters = pars,
                                points = 500, state.names = c("x", "y"))

  # draw trajectory
  #for (start in c( c(5,5), c(-5,-5))) {
  #  model.trajectory = trajectory(mod, y0 = c(start[1], start[2]), 
  #                                tlim = c(0, 10), 
  #                                parameters = pars, state.names = c("x", "y"))
  #}
  
  
  model.trajectory = trajectory(mod, y0 = c(-5, -5), 
                                tlim = c(0, 10), 
                                parameters = pars, state.names = c("x", "y"))
  model.trajectory = trajectory(mod, y0 = c(-5, 5), 
                                tlim = c(0, 10), 
                                parameters = pars, state.names = c("x", "y"))
  model.trajectory = trajectory(mod, y0 = c(5, -5), 
                                tlim = c(0, 10), 
                                parameters = pars, state.names = c("x", "y"))
  model.trajectory = trajectory(mod, y0 = c(5, 5), 
                                tlim = c(0, 10), 
                                parameters = pars, state.names = c("x", "y"))
  model.trajectory = trajectory(mod, y0 = c(5, 0), 
                                tlim = c(0, 10), 
                                parameters = pars, state.names = c("x", "y"))
  model.trajectory = trajectory(mod, y0 = c(-5, 0), 
                                tlim = c(0, 10), 
                                parameters = pars, state.names = c("x", "y"))
                                
  # Calculate the stability at the selected equilibrium
  stab = stability(mod, parameters = pars, system = "two.dim",
                 ystar=c(0, 0), state.names = c("x", "y"))
  print("")
  
}
```

# Question 10 NFkB oscillations with S = 0

```{r}
nfkb = function(timepoint, state, parameters) { 
  with(as.list(c(state, parameters)), {
    dI = -S + a * N - b * I 
    dN = -c * I - d * N 
    list(c(dI, dN))
  }) 
}

pars = c(a = 3, b = 1, c = 2, d = 1, S = 0)
pars = c(S = 0, a = 3, b = 1, c = 2, d = 1)

# for setting the axis limits
xlim = c(-10, 10) 
ylim = c(-15, 15)

inival = c(10, 10)
time_int = c(0, 20)

# draw vector field
model.flowField = flowField(nfkb, 
                            xlim = xlim, ylim = ylim, 
                            parameters = pars, 
                            state.names = c("I", "N"),
                            points = 19, add = FALSE, 
                            xlab = "I", ylab = "N",)

# draw nullclines
model.nullclines = nullclines(nfkb, xlim = xlim, ylim = ylim,
                              parameters = pars,
                              points = 500, state.names = c("I", "N"))

# draw trajectory
model.trajectory = trajectory(nfkb, y0 = c(-5, -5), tlim = c(0, 10), 
                              parameters = pars, state.names = c("I", "N"))

# Calulate the N and I over time
model.numericalSolution = numericalSolution(nfkb, 
                                            y0 = inival, 
                                            tlim = time_int, 
                                            type = "one",
                                            parameters = pars, 
                                            state.names = c("I", "N"),
                                            col = c("green", "orange"), 
                                            ylim = ylim, 
                                            add.legend = F)
legend("topright", col = c("green", "orange"), legend = c("x", "y"), lty = 1)

# Determine stability
stab = stability(nfkb, parameters = pars, system = "two.dim",
                 ystar = c(0, 0), state.names = c("I", "N"))

```

# Question 11 NFkB oscillations with S != 0

```{r}
nfkb = function(timepoint, state, parameters) { 
  with(as.list(c(state, parameters)), {
    dI = -S + a*N - b*I 
    dN = -c*I - d*N 
    list(c(dI, dN))
  }) 
}

pars = c(a = 3, b = 1, c = 2, d = 1, S = 5)
inival = c(-5, -5)
time_int = c(0, 20)

# for setting the axis limits
xlim = c(-10, 10) 
ylim = c(-5, 5)

# draw vector field
model.flowField = flowField(nfkb, 
                            xlim = xlim, ylim = ylim, 
                            parameters = pars,
                            state.names = c("I", "N"),
                            points = 19, 
                            add = FALSE, 
                            xlab = "I", 
                            ylab = "N", )

# draw nullclines
model.nullclines = nullclines(nfkb, 
                              xlim = xlim, ylim = ylim,
                              parameters = pars,
                              points = 500, 
                              state.names = c("I", "N"))

# draw trajectory

model.trajectory = trajectory(nfkb, 
                              y0 = inival, 
                              tlim = time_int, 
                              parameters = pars, 
                              state.names = c("I", "N"))

# Calulate the N and I over time
inival = c(0, 0)
model.numericalSolution = numericalSolution(nfkb, 
                                            y0 = inival, 
                                            tlim = time_int, 
                                            type = "one",
                                            parameters = pars, 
                                            state.names = c("I", "N"),
                                            col = c("green", "orange"), 
                                            ylim = ylim, 
                                            add.legend = F)
legend("topright", col = c("green", "orange"), legend = c("I", "N"), lty = 1)

# Plot x against y - why does this not work well?
# ggplot() +
#  geom_line(mapping = aes(x = model.numericalSolution$x, y = model.numericalSolution$y))

# Determine stability
stab = stability(nfkb, 
                 parameters = pars, 
                 system = "two.dim",
                 ystar=c(0, 0), 
                 state.names = c("I", "N"))

```
# Question 14

```{r}
# draw vector field
nfkb.flowField = flowField(xlab = "IkB", ylab = "NFkB", 
                           nfkb, 
                           xlim = xlim, ylim = ylim, 
                           parameters = pars, 
                           points = 19, 
                           add = FALSE, 
                           state.names = c("I", "N"))
# draw nullclines
nfkb.nullclines = nullclines(nfkb, 
                             xlim = xlim, ylim = ylim, 
                             parameters = pars, 
                             points = 500, 
                             state.names = c("I", "N"))
# draw trajectory
nfkb.trajectory = trajectory(nfkb, 
                             y0 = c(-5, -5), 
                             tlim = c(0, 100), 
                             parameters = pars, 
                             state.names = c("I", "N"))

# determine stability of equilibrium point
eq = c(tail(model.numericalSolution$x, 1), tail(model.numericalSolution$y, 1))
model.stability = stability(nfkb, 
                            ystar = eq, 
                            parameters = pars, 
                            state.names = c("I","N"))
```



# Question 16 Auto inhibition for NFkB and IkB aligned and change together

Not clear how to model this best

```{r}
pars_ori = c(a = 3, b = 1, c = 2, d = 1, S = 0)
pars_low = c(a = 3, b = 0.1, c = 2, d = 0.1, S = 0)
pars_high = c(a = 3, b = 10, c = 2, d = 10, S = 0)

pars = pars_low

# for setting the axis limits
xlim = c(-10, 10) 
ylim = c(-10, 10)

# draw vector field
model.flowField = flowField(nfkb, 
                            xlim = xlim, ylim = ylim, 
                            parameters = pars, 
                            state.names = c("I", "N"),
                            points = 19, add = FALSE, 
                            xlab = "I", ylab = "N")

# draw nullclines
model.nullclines = nullclines(nfkb, 
                              xlim = xlim, ylim = ylim,
                              parameters = pars,
                              points = 500, state.names = c("I", "N"))

# draw trajectory
model.trajectory = trajectory(nfkb, 
                              y0 = c(-5, -5), 
                              tlim = c(0, 100), 
                              parameters = pars, 
                              state.names = c("I", "N"))

# Calulate the N and I over time
model.numericalSolution = numericalSolution(nfkb, y0 = c(-5, -5), 
                                            tlim = c(0, 20), 
                                            type = "one",
                                            parameters = pars, 
                                            state.names = c("I", "N"),
                                            col = c("green", "orange"), 
                                            ylim = ylim, 
                                            add.legend = F)
legend("topright", col = c("green", "orange"), legend = c("x", "y"), lty = 1)

# Determine stability
stab = stability(nfkb, 
                 parameters = pars, 
                 system = "two.dim",
                 ystar=c(0, 0), 
                 state.names = c("I", "N"))
```


# Question 17 Auto inhibition removed

This means that parameters b and d become 0

```{r}
pars = c(a = 3, b = 0, c = 2, d = 0, S = 5)

# for setting the axis limits
xlim = c(-10, 10) 
ylim = c(-10, 10)

# draw vector field
model.flowField = flowField(nfkb, 
                            xlim = xlim, ylim = ylim, 
                            parameters = pars, 
                            state.names = c("I", "N"),
                            points = 19, add = FALSE, 
                            xlab = "I", ylab = "N")

# draw nullclines
model.nullclines = nullclines(nfkb, 
                              xlim = xlim, ylim = ylim,
                              parameters = pars,
                              points = 500, state.names = c("I", "N"))

# draw trajectory
model.trajectory = trajectory(nfkb, 
                              y0 = c(-5, -5), 
                              tlim = c(0, 100), 
                              parameters = pars, 
                              state.names = c("I", "N"))
model.trajectory = trajectory(nfkb, 
                              y0 = c(-2, -2), 
                              tlim = c(0, 100), 
                              parameters = pars, 
                              state.names = c("I", "N"))

# Calulate the N and I over time
model.numericalSolution = numericalSolution(nfkb, y0 = c(-5, -5), 
                                            tlim = c(0, 20), 
                                            type = "one",
                                            parameters = pars, 
                                            state.names = c("I", "N"),
                                            col = c("green", "orange"), 
                                            ylim = ylim, 
                                            add.legend = F)
legend("topright", col = c("green", "orange"), legend = c("x", "y"), lty = 1)

# Determine stability
stab = stability(nfkb, 
                 parameters = pars, 
                 system = "two.dim",
                 ystar=c(0, 0), 
                 state.names = c("I", "N"))
```


# Question 18 Negative Auto inhibition 

This means that parameters b and d become negative. You get immediate oscillating instability

```{r}
pars = c(a = 3, b = -1, c = 2, d = -1, S = 5)

# for setting the axis limits
xlim = c(-10, 10) 
ylim = c(-10, 10)

# draw vector field
model.flowField = flowField(nfkb, 
                            xlim = xlim, ylim = ylim, 
                            parameters = pars, 
                            points = 19, 
                            add = FALSE, 
                            state.names = c("I", "N"), xlab = "I", ylab = "N")

# draw nullclines
model.nullclines = nullclines(nfkb, 
                              xlim = xlim, ylim = ylim,
                              parameters = pars,
                              points = 500, 
                              state.names = c("I", "N"))

# draw trajectory
model.trajectory = trajectory(nfkb, 
                              y0 = c(0, 0), 
                              tlim = c(0, 10), 
                              parameters = pars, 
                              state.names = c("I", "N"))

# Calulate the N and I over time
ylim = c(-1000, 1000)
model.numericalSolution = numericalSolution(nfkb, 
                                            y0 = c(-5, -5), 
                                            tlim = c(0, 20), 
                                            type = "one",
                                            parameters = pars, 
                                            state.names = c("I", "N"),
                                            col = c("green", "orange"), 
                                            ylim = ylim, add.legend = F)
legend("topright", col = c("green", "orange"), legend = c("x", "y"), lty = 1)
```
