---
title: "Class 1 - The various types of stability and equilibrium plots in the slides"
author: "Hans"
date: "9/6/2019"
output: html_document
---


Single analysis
phasePlaneAnalysis(model, xlim = xlim, ylim = ylim, parameters = pars)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("phaseR")
#library(matlib)

EPSILON = 1e-10

are_same <- function (a, b) {
    return (abs(a - b) < EPSILON)
}
```


## Trace, det and eigenvalues
This function uses trace, det and eigen values to determine the type of equilibrium 

```{r}
test_shape <- function (e_values, trace, det) {
  ev1 = e_values[[1]]
  ev2 = e_values[[2]]
  
  if (det <= 0) {
    return('Saddle')
  }
  else if ( trace < 0 ) {   # Stable
    if (class(ev1) == "complex") {
      return ('Stable Spiral')
    }
    else {
      return ('Stable Node')
    }
  }
  else if (trace > 0) {  # Unstable
    if (class(ev1) == "complex") {
      return ('Unstable Spiral')
    }
    else {
      return ('Unstable Node')
    }
  }
  else {
    return('Centre')
  }
}
```


## Jacobian
This function uses just the eigen values to determine the type of equilibrium 
```{r}
test_jacobian <- function (Jacobian) {

  ev = eigen(Jacobian)

  ev1 = ev$values[[1]]
  ev2 = ev$values[[2]]
  
  ev1
  ev2

  if (are_same(Re(ev1),0)  && are_same(Re(ev1),0)) {
     return("Centre")   
  }
  else if (Re(ev1) < 0 && Re(ev2) < 0) { #stable
    if (class(ev1) == "complex") {
      return ('Stable Spiral')
    }
    else {
      return ('Stable Node')
    }
  }
  else if (Re(ev1) > 0 && Re(ev2) > 0) { #unstable
    if (class(ev1) == "complex") {
      return ('Unstable Spiral')
    }
    else {
      return ('Unstable Node')
    }
  }
}


tr <- function (Matr) {
  return(Matr[1,1] + Matr[2,2])
}

```


## Definition 
Here we start defining the various equations and there corresponding Jacobian
The name is indicative of the equlibrium type

```{r snode}
snode = function(timepoint, state, parameters) { 
  with(as.list(c(state, parameters)), {
    dx = -2*x - y 
    dy = -x - 2*y 
    list(c(dx, dy))
  }) 
}

(snode_jacobian = matrix(c (-2, -1, -1,-2), 2, 2, byrow=TRUE))

```


```{r unode}
unode = function(timepoint, state, parameters) { 
  with(as.list(c(state, parameters)), {
    dx = 2*x + y 
    dy = x + 2*y 
    list(c(dx, dy))
  }) 
}

(unode_jacobian = matrix(c (2, 1, 1,2), 2, 2, byrow=TRUE))
```


```{r saddle}
saddle = function(timepoint, state, parameters) { 
  with(as.list(c(state, parameters)), {
    dx = -x - y 
    dy = -2*x - y 
    list(c(dx, dy))
  }) 
}

(saddle_jacobian = matrix(c (-1, -1, -2, -1), 2, 2, byrow=TRUE))
```


```{r sspiral}
sspiral = function(timepoint, state, parameters) { 
  with(as.list(c(state, parameters)), {
    dx = -x + 2*y 
    dy = -2*x - y 
    list(c(dx, dy))
  }) 
}

(sspiral_jacobian = matrix(c (-1, 2, -2, -1), 2, 2, byrow=TRUE))
```


```{r uspiral}
uspiral = function(timepoint, state, parameters) { 
  with(as.list(c(state, parameters)), {
    dx = x + 2*y 
    dy = -2*x + y 
    list(c(dx, dy))
  }) 
}

(uspiral_jacobian = matrix(c (1, 2, -2, 1), 2, 2, byrow=TRUE))
```


```{r center}
center = function(timepoint, state, parameters) { 
  with(as.list(c(state, parameters)), {
    dx = x + 2*y 
    dy = -2*x - y 
    list(c(dx, dy))
  }) 
}
(center_jacobian = matrix(c (1, 2, -2, -1), 2, 2, byrow=TRUE))

```

# Here starts the analyis part

You need to select here one of the models you want to use:

```{r}
model = center
jac   = center_jacobian


```


```{r analyse}

# Show the model

jac

# Calculate the eigen values
ev = eigen(jac)
(values = ev$values)
(vectors = ev$vectors)

# Trace and Determinant 
trace = tr(jac)
det  = det(jac)

cat("The trace for this model is ", trace )
cat("The determinant for this model is ", det )

test_shape(values, trace, det)
test_jacobian(jac)

# for setting the axis limits
xlim = c(-10, 10) 
ylim = c(-10, 10)

# no parameters in the ODE, so we set pars as an empty vector
pars = vector()

# Calculate the stability at the selected equilibrium
stab = stability(model, parameters = pars, system = "two.dim",
                 ystar=c(0, 0), state.names = c("x", "y"))
 
# draw vector field
model_flowField = flowField(xlab = "x", ylab = "y", model, 
                            xlim = xlim, ylim = ylim, 
                            parameters = pars, 
                            points = 19, add = FALSE, 
                            state.names = c("x", "y"))
  
# draw nullclines
model_nullclines = nullclines(model, xlim = xlim, ylim = ylim,
                              parameters = pars,
                              points = 500, state.names = c("x", "y"))
  

  
model_trajectory = trajectory(model, y0 = c(-5, -5), 
                                tlim = c(0, 10), 
                                parameters = pars, state.names = c("x", "y"))
model_trajectory = trajectory(model, y0 = c(-5, 5), 
                                tlim = c(0, 10), 
                                parameters = pars, state.names = c("x", "y"))
model_trajectory = trajectory(model, y0 = c(5, -5), 
                                tlim = c(0, 10), 
                                parameters = pars, state.names = c("x", "y"))
model_trajectory = trajectory(model, y0 = c(5, 5), 
                                tlim = c(0, 10), 
                                parameters = pars, state.names = c("x", "y"))
model_trajectory = trajectory(model, y0 = c(5, 0), 
                                tlim = c(0, 10), 
                                parameters = pars, state.names = c("x", "y"))
model_trajectory = trajectory(model, y0 = c(-5, 0), 
                                tlim = c(0, 10), 
                                parameters = pars, state.names = c("x", "y"))
  

```




